---
title: "Method Received Anova"
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
getwd()

```

The purpose of this notebook is to do statistical analysis on the
dependent variable - average ticket completion time first with the
independent variable - submission method, then department.

Basically, we want to confirm that the difference in means that we have
demonstrated via Tableau are statistically signficant using ANOVA.
```{r}

```

Then we will use TUKEY HSD to determine which of those differences are
statistically significant as well.

```{r}
#importing sample data for Submission Method
method<-read.csv("C:/Users/Michael/Desktop/School/internship/method2.csv", header=T,sep =",")
```

```{r}
head(method)
```

```{r}
# EDA on sample data
table(method$method)
sort(tapply(method$days, method$method, mean))
```

```{r}
# Running analysis of variance
aov_meth = aov(days~method, data=method)
summary(aov_meth)
```

Our p-value is extremely low (\<2e-16) so the difference in the means is
statistically significant.

```{r}
# Now we will run the Tukey Honest Significance Difference Test to see which methods can statistically proven to be "better" than the others in terms of AVG ticket completion time
TukeyHSD(aov_meth)
```

The main columns we want to compare are: 
PHONE-CELL PHONE APPS -41.1562883 -45.367627 -36.944949 0.0000000 
PHONE-EMAIL -41.9395689 -47.731002 -36.148135 0.0000000

WEB-CELL PHONE APPS -29.6180947 -35.124043 -24.112147 0.0000000
WEB-EMAIL -30.4013753 -37.192604 -23.610147 0.0000000

WEB-PHONE 11.5381936 7.715854 15.360533 0.0000000

PHONE-INHOUSE 9.8582448 2.480468 17.236022 0.0019467

Our ANOVA proves that there is a statistically significant difference in
the means.

The Tukey test then proves that there is a statistically significant difference between the values of Android, Iphone, Email vs Web (and phone), as well as Web vs phone, and finally phone vs inhouse.

So we can say that the methods ranked from worst to best are: 
EMAIL/CELL (worst) 
PHONE APPS 
WEB 
PHONE 
INHOUSE (best)
```{r}
department<-read.csv("C:/Users/Michael/Desktop/School/internship/department.csv", header=T,sep =",")
```

```{r}
# inspecting our imported csv
table(department$dept)
```


```{r}
# looking at the sample means of the departments actual completion time
sort(tapply(department$days, department$dept, mean))
```

```{r}
# running anova on independent variable DEPARTMENT with ACTUAL COMPLETED DAYS as the dependent variable.
aov_dept = aov(days~dept, data=department)
summary(aov_dept)
```

```{r}
# now running TUKEY HSD test on the department data
TukeyHSD(aov_dept)
```

This data is a bit difficult to summarize so let's revisit the sample means but eliminate those that had too small of a sample size (\<.77% of total emergency tickets). Please note that due to sample size issues,
none of the departments eliminated had statistically adjusted p-values from TUKEY HSD.

DEPARTMENT                                MEAN COMPLETION TIME 
Enforcement Section                       6.72 
Solid Waste Management                    10.12 
PublicWorks: Road And Bridges             35.27 
Parks and Rec: RAAM                       49.01 
PublicWorks: Traffic Signals And Signs    61.66 
PublicWorks: Traffic Engineering          81.05

Now we want to look at the adjusted p-values for the comparisons we received via TUKEY HSD test. Let's pick out all the adjusted p-values that are extremely close to 0.

PublicWorks: Traffic Signals And Signs-PublicWorks: Traffic Engineering 0.0000000 
PublicWorks: Traffic Engineering-Parks and Rec: RAAM 0.0000000
PublicWorks: Traffic Engineering-Enforcement Section 0.0000000
PublicWorks: Traffic Engineering-PublicWorks: Road And Bridges 0.0000000
Solid Waste Management-PublicWorks: Traffic Engineering 0.0000000
Regulatory and Economic Resources-PublicWorks: Traffic Engineering 0.0001747 
Parks and Rec: RAAM-Enforcement Section 0.0000000 
Regulatory and Economic Resources-Enforcement Section 0.0025796 
PublicWorks: Road And Bridges-Enforcement Section 0.0000000 
PublicWorks: Road And Bridges-Parks and Rec: RAAM 0.0178070 
PublicWorks: Traffic Signals And Signs-Parks and Rec: RAAM 0.0000000 
PublicWorks: Traffic Signals And Signs-Enforcement Section 0.0000000 
PublicWorks: Traffic Signals And Signs-PublicWorks: Road And Bridges 0.0000000 
Solid Waste Management-Parks and Rec: RAAM 0.0000000 
Solid Waste Management-PublicWorks: Road And Bridges 0.0001151 
Solid Waste Management-PublicWorks: Traffic Signals And Signs 0.0000000

#### Now we will go from the down our list of sample means, as make conclusions.

**First, let's compare Enforcement Section and Solid Waste Management:**
Solid Waste Management-Enforcement Section 0.9984115

*This adjust p-value is too high, therefore we cannot say that there is a statistically significant difference between Solid Wast Managment and Enforcement Section*

**Next, let's compare Enforcement Section and Solid Waste Management to Public Works: Roads & Bridges:** PublicWorks: Road And Bridges-Enforcement Section 0.0000000 
Solid Waste Management-PublicWorks: Road And Bridges 0.0001151

*So we can conclude the difference (Enforcement Section and Solid Waste Management) and Public Works: Roads & Bridges is statistically significant because the adjusted P-values are extremely low.*

**Let's compare Public Works: Roads & Bridges to Parks and Rec: RAAM:**
PublicWorks: Road And Bridges-Parks and Rec: RAAM 0.0178070

*The difference between Public Works: Roads & Bridges to Parks and Rec: RAAM is statistically significant.*

**Let's compare Parks and Rec: RAAM: to Public Works: Traffic Signals And Signs:** 
PublicWorks: Traffic Signals And Signs-Parks and Rec: RAAM 0.0000000

*The difference between Parks and Rec: RAAM: to Public Works: Traffic Signals And Signs is statistically significant.*

**Finally, let's compare Public Works: Traffic Signals And Signs to Public Works: Traffic Engineering:** PublicWorks: Traffic Signals And Signs-PublicWorks: Traffic Engineering 0.0000000

*The difference between Public Works: Traffic Signals And Signs to Public Works: Traffic Engineering is statistically significant.*

So we can say that the Departments ranked from best to worst are (in terms of actual ticket completion time): Enforcement Section and Solid Waste Management (best) 
Public Works: Roads & Bridges 
Parks and Rec: RAAM 
Public Works: Traffic Signals And Signs 
Public Works: Traffic Engineering (worst)
